Fitting the Undetected: Light-curve Feature Estimation in the Presence of Non-detections

### 2012-04-14 tex/ms.tex

In addition to the expected brightness in Equation \ref{eq:fourier}, we assume that there is uncertainty or inappropriateness in the model, leading to a \emph{fractional model variance} $\eta^2$ at each point (assumed constant but easily generalized) as well as random variation due to photo arrival counts, assumed normal mean 0 with variance $\sigma_i^2$. The reported uncertainties $s_i^2$ are an estimate of the $\sigma_i^2$, see Subsection \ref{sec:uncertainty}. This leads to a true flux measurement of
\begin{eqnarray}\displaystyle
\ftrue = \mu_i + e_i
\label{eq:truefluxdef}
\end{eqnarray}
where $e_i \sim N(0,\sigma_i^2 + \eta^2\,\mu_i^2$. Recall that this is a latent variable because we will not see it when an observation is censored. We instantiate a latent variable, $b_i$, to represent the detection limit, in units of flux, at epoch $i$.  The $b_i$ parameter is essential because it probabilistically constrains the possible values of the mean brightness, $\mu_i$, of the light curve when there is a non-detection.  By employing a hierarchical model for the distribution of $b_i$, we can fully utilize all of the information encoded in both the detections and non-detections when computing the data likelihood.   At each epoch, we connect the observed flux, $\fobs$, to the latent variable $\ftrue$ by
\begin{eqnarray}\displaystyle
\fobs &=& \left\{\begin{array}{ll}
  \ftrue & \mbox{if $\ftrue \ge b_i$} \\
  \texttt{NA} & \mbox{if $\ftrue < b_i$} \\
\end{array} \right.
\end{eqnarray}
The observed flux is \texttt{NA} only when the true flux, $\ftrue$, is below the detection limit $b_i$.

@@

Finally if we assume that the data collected at each epoch are independent given the model parameters, we have that
\begin{eqnarray}\displaystyle
p(D|\theta,I) &=& \prod_i p(D_i|\theta,I)
\quad.
\end{eqnarray}
This is the likelihood for the entire
data set (all the measurements and non-detections of this star from
all the epochs, as delivered by the survey) given the $2K + 5$ parameter vector $\theta$.
Figure \ref{fig:graph} displays a graphical model highlighting dependencies and 
independence assumptions. This model
``correctly'' or at least ``justifiably'' uses all of the information
available, without making strong assumptions about the survey or its
veracity.

>>>>

In addition to the expected brightness in Equation \ref{eq:fourier}, we assume that there is uncertainty or inappropriateness in the model, leading to a \emph{fractional model variance} $\eta^2$ at each point (assumed constant but easily generalized) as well as random variation due to photon arrival counts, assumed normal mean 0 with variance $\sigma_i^2$. The reported uncertainties $s_i^2$ are an estimate of the $\sigma_i^2$, see Subsection \ref{sec:uncertainty}. This leads to a true flux measurement of
\begin{eqnarray}\displaystyle
\ftrue = \mu_i + e_i
\label{eq:truefluxdef}
\end{eqnarray}
where $e_i \sim N(0,\sigma_i^2 + \eta^2\,\mu_i^2)$. Recall that $f_i^*$ is a latent variable because we will not see it when an observation is censored. We instantiate a latent variable, $b_i$, to represent the detection limit, in units of flux, at epoch $i$.  The $b_i$ parameter is essential because it probabilistically constrains the possible values of the mean brightness, $\mu_i$, of the light curve when there is a non-detection.  By employing a hierarchical model for the distribution of $b_i$, we can fully utilize all of the information encoded in both the detections and non-detections when computing the data likelihood.   At each epoch, we connect the observed flux, $\fobs$, to the latent variable $\ftrue$ by
\begin{eqnarray}\displaystyle
\fobs &=& \left\{\begin{array}{ll}
  \ftrue & \mbox{if $\ftrue \ge b_i$} \\
  \texttt{NA} & \mbox{if $\ftrue < b_i$} \\
\end{array} \right.
\end{eqnarray}
The observed flux is \texttt{NA} only when the true flux, $\ftrue$, is below the detection limit $b_i$.

@@

Finally if we assume that the data collected at each epoch are independent given the model parameters, we have that
\begin{eqnarray}\displaystyle
p(D|\theta,I) &=& \prod_i p(D_i|\theta,I)
\quad.
\end{eqnarray}
This is the likelihood for the entire
data set (all the measurements and non-detections of this star from
all the epochs, as delivered by the survey) given the $2K + 9$ parameter vector $\theta$.
Figure \ref{fig:graph} displays a graphical model highlighting dependencies and 
independence assumptions. This model
``correctly'' or at least ``justifiably'' uses all of the information
available, without making strong assumptions about the survey or its
veracity.
